---
title: "Task1"
author: "Kirill Tumanov"
date: "18 09 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(dplyr)
library(reshape2)
```
##Load Data
```{r}
url1 <- "http://data.ssb.no/api/v0/dataset/95274.csv?lang=no"
url2 <- "http://data.ssb.no/api/v0/dataset/95276.csv?lang=no"
data_1 <- read.csv(url1,sep = ";",dec = ",")
data_2 <- read.csv(url2,sep = ";",dec = ",")
```
#Change names of column
```{r}
colnames <- c("region", "date", "variable", "value")
colnames(data_1) <- colnames
colnames(data_2) <- colnames
```
#Change the data
```{r}
data_1 <- data_1 %>%
  mutate(date = gsub("M","01",date)) %>%
  mutate (date = lubridate::ydm(date))
data_2 <- data_2 %>%
  mutate(date = gsub("M","01",date)) %>%
  mutate (date = lubridate::ydm(date))
```
#Recode the Norwegian labels into English
```{r}
data_1 <- data_1 %>%
  mutate(variable = recode (variable,
                          "Kapasitetsutnytting av senger (prosent)" = "Beds percentage capacity utilization",
                          "Kapasitetsutnytting av rom (prosent)" = "Rooms percentage capacity utilization",
                          "Pris per rom (kr)" = "Price per room" ))
data_2 <- data_2 %>%
  mutate(variable = recode (variable,
                          "Kapasitetsutnytting av senger (prosent)" = "Beds percentage capacity utilization",
                          "Kapasitetsutnytting av rom (prosent)" = "Rooms percentage capacity utilization",
                          "Pris per rom (kr)" = "Price per room" ))
data_1$value[data_1$value == 0] <- NA
data_2$value[data_2$value == 0] <- NA
```
#Join data
```{r}
data_1 <- data_1 %>%
  mutate (region = as.character(region),
         value = as.numeric(value))
data_2 <- data_2 %>%
  mutate (region = as.character(region),
         value = as.numeric(value))
total_data <- merge(data_1,data_2, by = c ("date","variable"))
```
#calculate
```{r}
total_data <- total_data %>%
  mutate(value_deduct = value.x - value.y)
means_region <- total_data %>%
  filter(variable == "Price per room") %>%
  group_by(region.x) %>%
  summarise(mean = mean(value_deduct, na.rm = TRUE))
means_years <- total_data %>%
  filter(variable == "Price per room") %>%
  group_by(lubridate::year(date)) %>%
  summarise(mean = mean(value_deduct,na.rm = TRUE))
colnames(means_region) <- c("region","mean")
colnames(means_years) <- c("year","mean")
max_vec <- means_years$mean==max(means_years$mean, na.rm=TRUE)
max_vec[is.na(max_vec)] <- FALSE
max_year <- means_years[max_vec,"year"]
min_vec <- means_years$mean==min(means_years$mean, na.rm=TRUE)
min_vec[is.na(min_vec)] <- FALSE
min_year <- means_years[min_vec,"year"]
max_reg <- means_region$mean==max(means_region$mean, na.rm=TRUE)
max_reg[is.na(max_reg)] <- FALSE
max_reg <- means_region[max_reg,"region"]
min_reg <- means_region$mean==min(means_region$mean, na.rm=TRUE)
min_reg[is.na(min_reg)] <- FALSE
min_reg <- means_region[min_reg,"region"]
```
2 dataframes of information were analyzed, they showed the difference between county average room price and the national average room price per month.
`r max_year` sorts years with their value and shows the year with the highest price.
`r min_year` sorts years with their value and shows the year with the lowest price.
`r max_reg` sorts regions with their value and shows the region with the highest price.
`r min_reg` sorts regions with their value and shows the region with the lowest price.

##Plot

```{r,echo=FALSE, warning = FALSE}
total_data %>%
  filter(variable=="Price per room") %>%
  filter(region.x=="19 Troms - Romsa") %>%
  ggplot(aes(x=date, y=value_deduct)) +
  geom_line(col="blue") + labs(x = "Date", y = "Price per Room compared to National Average", title = "Monthly price difference for Troms county from 1999 until today") +
    scale_x_date(breaks='3 years', 
               limits = c(as.Date("1999-01-01"),Sys.Date()))+
  theme(text = element_text(family = "Bookman"),
        title = element_text(color = "yellow",size = 12)) +
  theme(plot.background = element_rect (fill = "lightgreen"), panel.background = element_rect (fill = "lightgreen"), plot.margin = unit(c(5,10,5,10), units = "mm"))
```

The graph shows that room prices in Troms County were mostly below average after 1999. However, since 2016, prices began to change more sharply.


##Question
```{r}
dat_expand <- dcast(total_data,date + region.x ~ variable, value.var = "value.x")
names(dat_expand) [3:5] <- c ("roomCap","bedCap","priceRoom")
coler_county <- dat_expand %>%
  filter(date >= as.Date("2010-01-01")) %>%
  group_by(region.x) %>%
  summarise(cor = cor(x = roomCap, y=priceRoom,
                           use = "pairwise.complete.obs",
                           method = "pearson"))
```



